---
title: "Spatial mapping of single cells with differential stability genes"
author: "Hani Kim"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This script perform spatial mapping of single cells from the [scNMT mouse gastrulation data](https://www.nature.com/articles/s41586-019-1825-8) unto [spatially resolved embryo](https://pubmed.ncbi.nlm.nih.gov/31391582/) generated from GEO-seq. 

# Loading 

## Packages
```{r, message = FALSE}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(scran)
})
```

## Load spatial coordinates and scNMT data
```{r}
coordinate.E6.5 <- read.delim2("../../../Data/200521_geometry_E6.5.txt", header = TRUE, row.names = 1, sep=" ")
coordinate.E7.5 <- read.delim2("../../../Data/200521_geometry_E7.5.txt", header = TRUE, row.names = 1, sep=" ")
rownames(coordinate.E7.5) <- coordinate.E7.5[,1]
coordinate.E7.5 <- coordinate.E7.5[,-1]

marioni <- readRDS("../../../Data/SingleCellExperiment_genesymbol.rds")
```

## Source functions
```{r}
source("../../../R/Cepo.R") 
source("../../../R/doDE.R")
source("../../../R/doDD.R")
source("../../../R/utils.R")
source("../../../R/doSpatialMapping.R")
```

# Identify spatial zip codes
```{r}
spatial_E6.5_E7.5 <- df2numericMat(read.delim2("../../../Data/E5.5-E7.5.all.combine.exprs.txt", header = TRUE, sep="\t", row.names = 1))
spatial_E6.5 <- spatial_E6.5_E7.5[,grep("E6.5", colnames(spatial_E6.5_E7.5))]
spatial_E7.5 <- spatial_E6.5_E7.5[,grep("E7.5", colnames(spatial_E6.5_E7.5))]
colnames(spatial_E6.5) <- gsub("\\..+", "", colnames(spatial_E6.5))
colnames(spatial_E7.5) <- gsub("\\..+", "", colnames(spatial_E7.5))

hvg_E6.5 <- scran::getTopHVGs(scran::modelGeneVar(spatial_E6.5), n=5000)
hvg_E7.5 <- scran::getTopHVGs(scran::modelGeneVar(spatial_E7.5), n=5000)

spatial_E6.5.zipCode <- unique(unlist(consensusRank(spatial_E6.5[hvg_E6.5,])))
spatial_E7.5.zipCode <- unique(unlist(consensusRank(spatial_E7.5[hvg_E7.5,])))
```

# Prepare scNMT dataset
```{r}
metadata.marioni <- colData(marioni)
table(metadata.marioni$stage)

marioni_E6.5 <- logcounts(marioni[,metadata.marioni$stage %in% c("E6.5")])
colnames(marioni_E6.5) <- metadata.marioni$sample[metadata.marioni$stage %in% c("E6.5")]
marioni_E6.5.lineage.init <- metadata.marioni$lineage[metadata.marioni$stage %in% c("E6.5")]
marioni_E6.5 <- marioni_E6.5[,marioni_E6.5.lineage.init %in% c("Visceral_endoderm", "Epiblast", "Primitive_Streak")]
marioni_E6.5 <- marioni_E6.5[!is.na(rownames(marioni_E6.5)),]
marioni_E6.5.lineage <- marioni_E6.5.lineage.init[marioni_E6.5.lineage.init %in% c("Visceral_endoderm", "Epiblast", "Primitive_Streak")]
marioni_E6.5.lineage <- gsub("_", " ", marioni_E6.5.lineage)
rownames(marioni_E6.5) <- make.names(rownames(marioni_E6.5), unique = TRUE)

marioni_E7.5 <- logcounts(marioni[,metadata.marioni$stage %in% c("E7.5")])
del <- which(is.na(rownames(marioni_E7.5)))
marioni_E7.5 <- marioni_E7.5[-del,]
colnames(marioni_E7.5) <- metadata.marioni$sample[metadata.marioni$stage %in% c("E7.5")]
marioni_E7.5.lineage.init <- metadata.marioni$lineage[metadata.marioni$stage %in% c("E7.5")]
marioni_E7.5 <- marioni_E7.5[,marioni_E7.5.lineage.init %in% c("Ectoderm", "Endoderm", "Mesoderm")]
marioni_E7.5.lineage <- marioni_E7.5.lineage.init[marioni_E7.5.lineage.init %in% c("Ectoderm", "Endoderm", "Mesoderm")]
rownames(marioni_E7.5) <- make.names(rownames(marioni_E7.5), unique = TRUE)
```

# Run Dx and perform spatial mapping of E6.5 single cells
```{r}
DS.list <- Cepo(marioni_E6.5, marioni_E6.5.lineage)
MAST.list <- doMAST(marioni_E6.5, marioni_E6.5.lineage)
Voom.list <- doVoom(marioni_E6.5, as.factor(marioni_E6.5.lineage))
DD.list <- doDD(marioni_E6.5, marioni_E6.5.lineage)

top <- 100
DS <- unique(unlist(lapply(DS.list, function(x)names(x)[1:top])))
MAST <- unique(unlist(lapply(MAST.list, function(x){rownames(x[order(x[,"pval"], decreasing = FALSE),])[1:top]})))
Voom <- unique(unlist(lapply(Voom.list, function(x){rownames(x[order(x[,"t"], decreasing = TRUE),])[1:top]})))
DD <- unique(unlist(lapply(DD.list, function(x){rownames(x[order(x[,1], decreasing = TRUE),])[1:top]})))

DS.map.E6.5 <- cbind(doSpatialMap(marioni_E6.5, spatial_E6.5, intersect(DS, rownames(spatial_E6.5)), coordinate = coordinate.E6.5, k=1), marioni_E6.5.lineage)
barplot(table(paste(DS.map.E6.5$labels, DS.map.E6.5$spatialLabel)), las=2, cex.names = 0.6)
t1 <- gsub("Epi", "Epiblast", gsub("PS", "Primitive Streak", gsub("E[12]", "Visceral endoderm", as.character(DS.map.E6.5$labels))))
sum(t1 == DS.map.E6.5$marioni_E6.5.lineage) / length(t1)

MAST.map.E6.5 <- cbind(doSpatialMap(marioni_E6.5, spatial_E6.5, intersect(MAST, rownames(spatial_E6.5)), coordinate = coordinate.E6.5, k=1), marioni_E6.5.lineage)
barplot(table(paste(MAST.map.E6.5$labels, MAST.map.E6.5$spatialLabel)), las=2, cex.names = 0.6)
t1 <- gsub("Epi", "Epiblast", gsub("PS", "Primitive Streak", gsub("E[12]", "Visceral endoderm", as.character(MAST.map.E6.5$labels))))
sum(t1 == MAST.map.E6.5$marioni_E6.5.lineage) / length(t1)

Voom.map.E6.5 <- cbind(doSpatialMap(marioni_E6.5, spatial_E6.5, intersect(Voom, rownames(spatial_E6.5)), coordinate = coordinate.E6.5, k=1), marioni_E6.5.lineage)
barplot(table(paste(Voom.map.E6.5$labels, Voom.map.E6.5$spatialLabel)), las=2, cex.names = 0.6)
t1 <- gsub("Epi", "Epiblast", gsub("PS", "Primitive Streak", gsub("E[12]", "Visceral endoderm", as.character(Voom.map.E6.5$labels))))
sum(t1 == Voom.map.E6.5$marioni_E6.5.lineage) / length(t1)

DD.map.E6.5 <- cbind(doSpatialMap(marioni_E6.5, spatial_E6.5, intersect(DD, rownames(spatial_E6.5)), coordinate = coordinate.E6.5, k=1), marioni_E6.5.lineage)
barplot(table(paste(DD.map.E6.5$labels, DD.map.E6.5$spatialLabel)), las=2, cex.names = 0.6)
t1 <- gsub("Epi", "Epiblast", gsub("PS", "Primitive Streak", gsub("E[12]", "Visceral endoderm", as.character(DD.map.E6.5$labels))))
sum(t1 == DD.map.E6.5$marioni_E6.5.lineage) / length(t1)
```

## Calculate purity of labels at spatial locations
```{r}
locs <- colnames(spatial_E6.5)
purityMat <- matrix(NA, nrow=length(locs), ncol=4)
colnames(purityMat) <- c("DS", "MAST", "Voom", "DD")
rownames(purityMat) <- locs

map1 <- split(DS.map.E6.5$marioni_E6.5.lineage, DS.map.E6.5$spatialLabel)
purityMat <- purity(map1, 1, purityMat, stage = "E6.5")
map2 <- split(MAST.map.E6.5$marioni_E6.5.lineage, MAST.map.E6.5$spatialLabel)
purityMat <- purity(map2, 2, purityMat, stage = "E6.5")
map3 <- split(Voom.map.E6.5$marioni_E6.5.lineage, Voom.map.E6.5$spatialLabel)
purityMat <- purity(map3, 3, purityMat, stage = "E6.5")
map4 <- split(DD.map.E6.5$marioni_E6.5.lineage, DD.map.E6.5$spatialLabel)
purityMat <- purity(map4, 4, purityMat, stage = "E6.5")

colSums(purityMat, na.rm = TRUE)
```

# Run Dx and perform spatial mapping of E7.5 single cells
```{r}
DS.list <- Cepo(marioni_E7.5, marioni_E7.5.lineage)
MAST.list <- doMAST(marioni_E7.5, marioni_E7.5.lineage)
Voom.list <- doVoom(marioni_E7.5, as.factor(marioni_E7.5.lineage))
DD.list <- doDD(marioni_E7.5, marioni_E7.5.lineage)

top <- 100
DS <- unique(unlist(lapply(DS.list, function(x)names(x)[1:top])))
MAST <- unique(unlist(lapply(MAST.list, function(x){rownames(x[order(x[,"pval"], decreasing = FALSE),])[1:top]})))
Voom <- unique(unlist(lapply(Voom.list, function(x){rownames(x[order(x[,"t"], decreasing = TRUE),])[1:top]})))
DD <- unique(unlist(lapply(DD.list, function(x){rownames(x[order(x[,1], decreasing = TRUE),])[1:top]})))


par(mfrow=c(2,2))
DS.map.E7.5 <- cbind(doSpatialMap(marioni_E7.5, spatial_E7.5, intersect(DS, rownames(spatial_E7.5)), coordinate = coordinate.E7.5, k=1), marioni_E7.5.lineage)
barplot(table(paste(DS.map.E7.5$labels, DS.map.E7.5$spatialLabel)), las=2, cex.names = 0.6, main="DS")
DS.Level1 <- gsub("^Ect$", "Ectoderm", gsub("^E$", "Endoderm", gsub("MP", "Mesoderm", gsub("PS", "Mesoderm", gsub("MA", "Mesoderm", gsub("[123]", "", DS.map.E7.5$labels))))))
sum(DS.Level1 == DS.map.E7.5$marioni_E7.5.lineage) / length(DS.Level1)

MAST.map.E7.5 <- cbind(doSpatialMap(marioni_E7.5, spatial_E7.5, intersect(MAST, rownames(spatial_E7.5)), coordinate = coordinate.E7.5, k=1), marioni_E7.5.lineage)
barplot(table(paste(MAST.map.E7.5$labels, MAST.map.E7.5$spatialLabel)), las=2, cex.names = 0.6, main="MAST")
MAST.Level1 <- gsub("^Ect$", "Ectoderm", gsub("^E$", "Endoderm", gsub("MP", "Mesoderm", gsub("PS", "Mesoderm", gsub("MA", "Mesoderm", gsub("[123]", "", MAST.map.E7.5$labels))))))
sum(MAST.Level1 == MAST.map.E7.5$marioni_E7.5.lineage) / length(MAST.Level1)

Voom.map.E7.5 <- cbind(doSpatialMap(marioni_E7.5, spatial_E7.5, intersect(Voom, rownames(spatial_E7.5)), coordinate = coordinate.E7.5, k=1), marioni_E7.5.lineage)
barplot(table(paste(Voom.map.E7.5$labels, Voom.map.E7.5$spatialLabel)), las=2, cex.names = 0.6, main="VoomS")
Voom.Level1 <- gsub("^Ect$", "Ectoderm", gsub("^E$", "Endoderm", gsub("MP", "Mesoderm", gsub("PS", "Mesoderm", gsub("MA", "Mesoderm", gsub("[123]", "", Voom.map.E7.5$labels))))))
sum(Voom.Level1 == Voom.map.E7.5$marioni_E7.5.lineage) / length(Voom.Level1)

DD.map.E7.5 <- cbind(doSpatialMap(marioni_E7.5, spatial_E7.5, intersect(DD, rownames(spatial_E7.5)), coordinate = coordinate.E7.5, k=1), marioni_E7.5.lineage)
barplot(table(paste(DD.map.E7.5$labels, DD.map.E7.5$spatialLabel)), las=2, cex.names = 0.6, main="DD")
DD.Level1 <- gsub("^Ect$", "Ectoderm", gsub("^E$", "Endoderm", gsub("MP", "Mesoderm", gsub("PS", "Mesoderm", gsub("MA", "Mesoderm", gsub("[123]", "", DD.map.E7.5$labels))))))
sum(DD.Level1 == DD.map.E7.5$marioni_E7.5.lineage) / length(DD.Level1)

names(Level1) <- colnames(marioni_E7.5)
Level2 <- map.E7.5.v2$labels
Level3 <- map.E7.5.v2$spatialLabel
length(colnames(spatial_E7.5))
```

## Calculate purity of labels at spatial locations
```{r}
locs <- colnames(spatial_E7.5)
purityMat <- matrix(NA, nrow=length(locs), ncol=4)
colnames(purityMat) <- c("DS", "MAST", "Voom", "DD")
rownames(purityMat) <- locs

map1 <- split(DS.map.E7.5$marioni_E7.5.lineage, DS.map.E7.5$spatialLabel)
purityMat <- purity(map1, 1, purityMat, stage = "E7.5")
map2 <- split(MAST.map.E7.5$marioni_E7.5.lineage, MAST.map.E7.5$spatialLabel)
purityMat <- purity(map2, 2, purityMat, stage = "E7.5")
map3 <- split(Voom.map.E7.5$marioni_E7.5.lineage, Voom.map.E7.5$spatialLabel)
purityMat <- purity(map3, 3, purityMat, stage = "E7.5")
map4 <- split(DD.map.E7.5$marioni_E7.5.lineage, DD.map.E7.5$spatialLabel)
purityMat <- purity(map4, 4, purityMat, stage = "E7.5")

pieMat.E7.5 <- purityMat

colSums(purityMat, na.rm = TRUE)
```

# Session Info
```{r}
sessioninfo::session_info()
```




