---
title: "Visualisation of spatial mapping results of single cells with differential stability genes"
author: "Hani Kim"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This script provide code to generate the results from spatial mapping of single cells from the [scNMT mouse gastrulation data](https://www.nature.com/articles/s41586-019-1825-8) unto [spatially resolved embryo](https://pubmed.ncbi.nlm.nih.gov/31391582/) generated from GEO-seq. Each spatial location is depicted as a pie chart, where the proportion of cell type labels mapped to locations are shown. 

# Loading 

## Packages
```{r, message = FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(dplyr)
})
```

## Load spatial mapping results
```{r}
load("../../../Data/spatial_pieMat.E6.5.RData", verbose=T)
load("../../../Data/spatial_pieMat.E7.5.RData", verbose=T)
```

# Generate results as pie charts for E6.5
```{r}
pieMat <- pieMat.E6.5
labels <- c("Epiblast", "Endoderm", "Mesoderm")

mat <- lapply(1:ncol(pieMat), function(x) {
    t <- do.call(rbind, lapply(pieMat[,x], function(y) y))
    t <- lapply(1:nrow(t), function(y) {
        if(is.na(t[y,])) {
            list(rep("0", 4))
        } else {
            strsplit(t[y,], " ")
        }
    })
    t <- do.call(rbind, lapply(t, function(x) as.numeric(x[[1]])))
    colnames(t) <- c(labels, "cell_number")
    rownames(t) <- rownames(pieMat)
    return(t)
})
names(mat) <- colnames(pieMat)

counter <- 0
mat_count <- lapply(mat, function(x) {
    counter <<- counter + 1
    method <- names(mat)[[counter]]
    t <- t(apply(x, 1, function(y) {
        y[1:3]
    }))
    t <- as.data.frame(t)
    tbl <- reshape2::melt(t)
    tbl$method <- factor(method)
    tbl$loc <- factor(rep(rownames(t), 3))
    tbl$size_raw <- as.numeric(rep(x[,"cell_number"], 3))
    tbl$size <- round(log(tbl$size_raw+1))
    tbl$size_2n <- tbl$size/2
    return(tbl)
})
mat_count <- do.call(rbind, mat_count)
mat_count$size <- scales::rescale(tbl$size, to=c(1,8))

lapply(sort(unique(mat_count$size_2n)), function(i) {
  table(mat_count[mat_count$size_2n == i,"size_raw"])  
})

g <- ggplot(mat_count, aes(x=size/2, y=value, fill=variable, width=size)) +
    geom_bar(position="fill", stat="identity") + coord_polar("y") + 
    facet_grid(method~loc) + 
    theme_classic() 
g %>% ggsave(file="../../../Figures/spatial/piechart_E7.5.pdf", width=40, height=8)
```

# Generate results as pie charts for E7.5
```{r}
pieMat <- pieMat.E7.5
labels <- c("Ectoderm", "Endoderm", "Mesoderm") #E7.5

mat <- lapply(1:ncol(pieMat), function(x) {
    t <- do.call(rbind, lapply(pieMat[,x], function(y) y))
    t <- lapply(1:nrow(t), function(y) {
        if(is.na(t[y,])) {
            list(rep("0", 4))
        } else {
            strsplit(t[y,], " ")
        }
    })
    t <- do.call(rbind, lapply(t, function(x) as.numeric(x[[1]])))
    colnames(t) <- c(labels, "cell_number")
    rownames(t) <- rownames(pieMat)
    return(t)
})
names(mat) <- colnames(pieMat)

counter <- 0
mat_count <- lapply(mat, function(x) {
    counter <<- counter + 1
    method <- names(mat)[[counter]]
    t <- t(apply(x, 1, function(y) {
        y[1:3]
    }))
    t <- as.data.frame(t)
    tbl <- reshape2::melt(t)
    tbl$method <- factor(method)
    tbl$loc <- factor(rep(rownames(t), 3))
    tbl$size_raw <- as.numeric(rep(x[,"cell_number"], 3))
    tbl$size <- round(log(tbl$size_raw+1))
    tbl$size_2n <- tbl$size/2
    return(tbl)
})
mat_count <- do.call(rbind, mat_count)
mat_count$size <- scales::rescale(tbl$size, to=c(1,8))

lapply(sort(unique(mat_count$size_2n)), function(i) {
  table(mat_count[mat_count$size_2n == i,"size_raw"])  
})

g <- ggplot(mat_count, aes(x=size/2, y=value, fill=variable, width=size)) +
    geom_bar(position="fill", stat="identity") + coord_polar("y") + 
    facet_grid(method~loc) + 
    theme_classic() 
g %>% ggsave(file="../../../Figures/spatial/piechart_E7.5.pdf", width=40, height=8)
```

# Session Info
```{r}
sessioninfo::session_info()
```