---
title: "Pairwise scatter plot of all simulation results"
author: "Hani Kim and Kevin Wang"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This script generates pairwise scatter plot of the differential analysis methods on the simulation datasets.

# Loading 

## Packages
```{r, message = FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(SingleCellExperiment)
  library(yardstick)
  library(ggpubr)
  library(RColorBrewer)
  library(ggplot2)
  library(ggthemes)
  library(dplyr)
  library(GGally)
})
```

## Source functions
```{r}
source("../../../R/simulateDS.R")
source("../../../R/utils.R")
```

## Load simulated scRNA-seq datasets
```{r}
load("../../../Data/201210_simulation_stability_genes.RData", verbose=T)
load("../../../Data/201210_simulation_params.RData", verbose=T)

load("../../../Data/simulation_dx_results_201210.RData", verbose=T)
load("../../../Data/simulation_de_results_201210.RData", verbose=T)
```

# Prepare Dx results
```{r}
sim_dx_result <- list(
  Cepo=sim_ds_res,
  Voom=sim_Voom_res,
  Limma=sim_Limma_res,
  ttest=sim_ttest_res,
  MAST=sim_MAST_res,
  Wilcoxon=sim_Wilcoxon_res,
  DD=sim_DD_res
)
```

# Calculate zprop, CV, and mean for each simulation dataset
```{r}
stats.gene <- lapply(sim_list, function(x) {
  mnz <- matrixStats::rowMeans2(x)
  names(mnz) <- names(x)
  sds <- matrixStats::rowSds(x)
  cv <- sds/mnz
  names(cv) <- names(x)
  zprop <- apply(x, 1, function(y) sum(y==0)/length(y))
  return(stats=list(mean=mnz, cv=cv, zprop=zprop))
})
```

# Generate dataframe of results for each simulation dataset
```{r}
ggsim_ranked <- lapply(1:49, function(i) {
  
  idx <- rownames(sim_ds)
  df=data.frame(Cepo=rank(sim_dx_result[["Cepo"]][[i]]$Group1)[idx],
                Voom=rank(sim_dx_result[["Voom"]][[i]]$Group1)[idx],
                Limma=rank(sim_dx_result[["Limma"]][[i]]$Group1)[idx],
                ttest=rank(sim_dx_result[["ttest"]][[i]]$Group1)[idx],
                Wilcoxon=rank(sim_dx_result[["Wilcoxon"]][[i]]$Group1)[idx],
                MAST=rank(sim_dx_result[["MAST"]][[i]]$Group1)[idx],
                DD=rank(sim_dx_result[["DD"]][[i]]$Group1)[idx],
                cv=rank(stats.gene[[i]][["cv"]])[idx],
                zprop=rank(stats.gene[[i]][["zprop"]])[idx],
                mean=rank(stats.gene[[i]][["mean"]])[idx]
  )
  rownames(df) = idx
  return(df)
  
})

ggsim_score <- lapply(1:49, function(i) {
  
  idx <- rownames(sim_ds)
  df=data.frame(Cepo=sim_dx_result[["Cepo"]][[i]]$Group1[idx],
                Voom=sim_dx_result[["Voom"]][[i]]$Group1[idx],
                Limma=sim_dx_result[["Limma"]][[i]]$Group1[idx],
                ttest=sim_dx_result[["ttest"]][[i]]$Group1[idx],
                Wilcoxon=sim_dx_result[["Wilcoxon"]][[i]]$Group1[idx],
                MAST=sim_dx_result[["MAST"]][[i]]$Group1[idx],
                DD=sim_dx_result[["DD"]][[i]]$Group1[idx],
                cv=stats.gene[[i]][["cv"]][idx],
                zprop=stats.gene[[i]][["zprop"]][idx],
                mean=stats.gene[[i]][["mean"]][idx]
  )
  rownames(df) = idx
  return(df)
  
})
```

# Plot pairwise results
```{r}
lapply(1:49, function(i) {
  
  type <- i
  GGally::ggpairs(ggsim_ranked[[i]],
                  1:7,
                  lower = list(continuous = wrap(GGscatterPlot, color_vec=ggsim_ranked[[i]]$zprop)),
                  upper = list(continuous = wrap("cor", method= "spearman"))) %>%
    ggsave(filename=paste0("../../../Figures/simulation/qc/simulation_", i,".pdf"), width=6, height=6)
  
})
```

# Session Info
```{r}
sessioninfo::session_info()
```



