---
title: "Cellbench scatter plot of the differential CV and zero proportions"
author: "Hani Kim"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This script generates scatter plot of coefficient of variation (CV) and proportion of zeros (zprop) against the ranks of the top 50 genes identified for each cell type in the 3-class Cellbench data. To illustrate the differential pattern, CV and zprop of the non-anchor cell types have been plotted on the same plot. 

# Loading 

## Packages
```{r, message = FALSE}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(GGally)
  library(RColorBrewer)
  library(ggplot2)
  library(ggthemes)
  library(dplyr)
})
```

## Load 3-class cellbench data and dx results
```{r}
load("../../../Data/cellbench_sce_3class.RData", verbose=T)
rm(sce_CELseq2_log)
rm(sce_Dropseq_log)

load("../../../Data/cellbench_DE_benchmark_results.RData", verbose=T)
load("../../../Data/cellbench_Dx_benchmark_results.RData", verbose=T)
```

## Source functions
```{r}
source("../../../R/utils.R")
```

# Calculate the two stability components of Cepo
```{r}
exprsMat <- logcounts(sce_10x_log)
celltype <- factor(sce_10x_log$celltype)

zprop <- apply(exprsMat, 1, function(x) length(x[x==0])/length(x))
cv <- apply(exprsMat, 1, function(x) sd(x)/mean(x))

stats_cty <- lapply(levels(celltype), function(x) {
  m <- exprsMat[,celltype %in% x]
  zprop_cty <- apply(m, 1, function(x) length(x[x==0])/length(x))
  cv_cty <- apply(m, 1, function(x) sd(x)/mean(x))
  zprop_cty <- zprop_cty[!is.na(cv_cty)]
  cv_cty <- cv_cty[!is.na(cv_cty)]
  return(list(zprop=zprop_cty, cv=cv_cty))
})
names(stats_cty) <- levels(celltype)
```

# Plot scatter of rank vs components
```{r fig.height=15, fig.width=20, message=FALSE}
dx.res <- list(Cepo=DS.res, Voom=Voom.res, DD=DD.res, mast=MAST.res)

dftoPlot <- lapply(levels(celltype), function(i) {
  
  idx <- names(stats_cty[[i]]$zprop)
  
  dx_mat <- lapply(seq_along(dx.res), function(dx) {
    dx_name <- names(dx.res)[[dx]]
    anchor <- i
    non_anchor <- setdiff(names(dx.res[[1]]), i)
    
    stats_anchr_df <- data.frame(
      rank=rank(-dx.res[[dx_name]][[i]])[idx],
      cv=stats_cty[[anchor]]$cv[idx],
      zprop=stats_cty[[anchor]]$zprop[idx],
      method=dx_name,
      anchor=TRUE,
      cty_major=i,
      cty_minor=i
    )
    rownames(stats_anchr_df) <- idx
    
    stats_nanchr_df <- lapply(non_anchor, function(x) {
      df <- data.frame(
        rank=rank(-dx.res[[dx_name]][[i]])[idx],
        cv=stats_cty[[x]]$cv[idx],
        zprop=stats_cty[[x]]$zprop[idx],
        method=dx_name,
        anchor=FALSE,
        cty_major=i,
        cty_minor=x
      )
      rownames(df) <- idx
      return(df)
    })
    stats_nanchr_df <- do.call(rbind, stats_nanchr_df)
    
    stats_df <- rbind(stats_anchr_df,
                      stats_nanchr_df)
    
    return(stats_df)
  })
  
  dx_mat <- do.call(rbind, dx_mat)
  return(dx_mat)
  
})
dftoPlot <- do.call(rbind, dftoPlot)
dftoPlot$method <- factor(dftoPlot$method, levels=c("Cepo", "Voom", "DD", "mast"))
dftoPlot <- dftoPlot[dftoPlot$rank < 51, ]

g1 <- ggplot(dftoPlot, aes(x=rank, y=cv, col=cty_minor)) + 
  geom_point() + geom_line(aes(group = rank)) +
  facet_wrap(~cty_major+method,ncol=4, scales = "free_y") + 
  coord_cartesian(xlim=c(0,50)) + theme_classic()

g2 <- ggplot(dftoPlot, aes(x=rank, y=zprop, col=cty_minor)) + 
  geom_point() + geom_line(aes(group = rank)) +
  facet_wrap(~cty_major+method,ncol=4) + 
  coord_cartesian(xlim=c(0,50)) + theme_classic()

ggpubr::ggarrange(g1,g2, ncol=1) %>% ggsave(filename="../../../Figures/cellbench/cellbench_components.pdf", width=20, height=15)
```

Generate boxplots
```{r}

dftoPlot <- lapply(levels(celltype), function(i) {
  
  idx <- names(stats_cty[[i]]$zprop)
  
  dx_mat <- lapply(seq_along(dx.res), function(dx) {
    dx_name <- names(dx.res)[[dx]]
    anchor <- i
    non_anchor <- setdiff(names(dx.res[[1]]), i)
    
    stats_anchr_df <- data.frame(
      rank=rank(-dx.res[[dx_name]][[i]])[idx],
      cv=stats_cty[[anchor]]$cv[idx],
      zprop=stats_cty[[anchor]]$zprop[idx],
      method=dx_name,
      anchor=TRUE,
      cty_major=i,
      cty_minor=i
    )
    rownames(stats_anchr_df) <- idx
    
    stats_nanchr_df <- lapply(non_anchor, function(x) {
      df <- data.frame(
        rank=rank(-dx.res[[dx_name]][[i]])[idx],
        cv=stats_cty[[x]]$cv[idx],
        zprop=stats_cty[[x]]$zprop[idx],
        method=dx_name,
        anchor=FALSE,
        cty_major=i,
        cty_minor=x
      )
      rownames(df) <- idx
      return(df)
    })
    
    diff.list <- lapply(stats_nanchr_df, function(x) {
      diff_cv = x$cv - stats_anchr_df$cv
      names(diff_cv) <- idx
      diff_zprop =  x$zprop - stats_anchr_df$zprop
      names(diff_zprop) <- idx
      return(list(cv=diff_cv, zprop=diff_zprop))
    })
    diff.sum <- lapply(c("cv", "zprop"), function(y) {
      rowMeans(do.call(cbind, lapply(diff.list, function(x) x[[y]])))
    })
    names(diff.sum) <- c("cv", "zprop")
    diff.mat <- as.data.frame(do.call(cbind, diff.sum))
    diff.mat$method = dx_name
    diff.mat$cty_major = i
    diff.mat$rank = stats_anchr_df$rank
  
    return(diff.mat)
  })
  
  dx_mat <- do.call(rbind, dx_mat)
  return(dx_mat)
  
})
dftoPlot <- do.call(rbind, dftoPlot)
dftoPlot$method <- factor(dftoPlot$method, levels=c("Cepo", "Voom", "DD", "mast"))
dftoBoxplot <- dftoPlot[dftoPlot$rank < 500, ]

g1 <- ggplot(dftoBoxplot, aes(x=method, y=cv, fill=method)) + 
  geom_boxplot(notch=F) + facet_wrap(~cty_major,ncol=4) + theme_classic()
g2 <- ggplot(dftoBoxplot, aes(x=method, y=zprop, fill=method)) + 
  geom_boxplot(notch=F) + facet_wrap(~cty_major,ncol=4) + theme_classic()

patchwork::wrap_plots(g1,g2,ncol=1) %>% ggsave(filename="../../../Figures/cellbench/cellbench_components_boxplot.pdf", width=5, height=10)

idx_cepo=dftoBoxplot$method == "Cepo"
idx_voom=dftoBoxplot$method == "Voom"
idx_dd=dftoBoxplot$method == "DD"
idx_mast=dftoBoxplot$method == "mast"

lapply(c("cv", "zprop"), function(i) {
  print(i)
  print(t.test(dftoBoxplot[,i][idx_cepo], dftoBoxplot[,i][idx_voom]))
  print(t.test(dftoBoxplot[,i][idx_cepo], dftoBoxplot[,i][idx_dd]))
  print(t.test(dftoBoxplot[,i][idx_cepo], dftoBoxplot[,i][idx_mast]))
})

```

# Session Info
```{r}
sessioninfo::session_info()
```

