---
title: "Trajectory analysis of HSPC differentiation using refined gene sets"
author: "Hani Kim"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This script runs trajectory analysis on the HSPC differentiation dataset using [dynverse](https://dynverse.org/). After features have been identified from running our benchmarking differential analysis algorithms, we find features that have been found exclusively by each of the differential analysis methods. We combine these exclusive gene sets with the common gene set (genes identified as top 200 genes by all four benchmarked methods).

# Loading 

## Packages
```{r, message = FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(SingleCellExperiment)
  library(hdf5r)
  library(dyno)
  library(scater)
  library(ggplot2)
  library(tidyverse)
  library(dyntoy)
  library(dyneval)
  library(patchwork)
})

#dynwrap::test_docker_installation(detailed = TRUE)
```

## Load HSPC single cell data and Dx results
```{r}
load("../../../Data/Dx_benchmark_result_HSPC.RData", verbose = T)
load("../../../Data/trajectory_sce_HSC.RData", verbose=T)
```

## Source functions
```{r}
source("../../../R/Cepo.R") 
source("../../../R/doDE.R")
source("../../../R/doDD.R")
source("../../../R/utils.R")
```

# Generate method-exclusive and common gene sets
```{r}
n=200
features_new <- lapply(names(DS.res), function(cty) {
  
  f1 = names(DS.res[[cty]][1:n])
  f2 = names(Voom.res[[cty]][1:n])
  f3 = names(DD.res[[cty]][1:n])
  f4 = names(MAST.res[[cty]][1:n])
  f_all <- list(cepo=f1,voom=f2,dd=f3,mast=f4)

  common <- Reduce(intersect, f_all)
  exclude_cepo <- setdiff(f1, unique(unlist(list(voom=f2,dd=f3,mast=f4))))
  exclude_voom <- setdiff(f2, unique(unlist(list(cepo=f1,dd=f3,mast=f4))))
  exclude_dd <- setdiff(f3, unique(unlist(list(cepo=f1,voom=f2,mast=f4))))
  exclude_mast <- setdiff(f4, unique(unlist(list(cepo=f1,voom=f2,dd=f3))))

  included <- list(cepo=c(common, exclude_cepo), 
                   voom=c(common, exclude_voom),
                   dd=c(common, exclude_dd),
                   mast=c(common, exclude_mast))
  
  return(included)
  
})
names(features_new) <- names(DS.res)

features_final <- lapply(1:4, function(y) {
  tmp <- lapply(1:8, function(x) {
    features_new[[x]][[y]]
  })
  tmp <- unique(unlist(tmp))
  return(tmp)
})
names(features_final) <- c("cepo", "voom", "dd", "mast")
```

# Visualise overlap of gene sets using Upset plots
```{r}
features_celltype <- lapply(1:4, function(y) {
  tmp <- lapply(1:8, function(x) {
    features_new[[x]][[y]]
  })
  names(tmp) <- names(DS.res)
  return(tmp)
})
names(features_celltype) <- c("cepo", "voom", "dd", "mast")

all.list <- list(cepo=lapply(DS.res, function(x) names(x[1:200])),
                 voom=lapply(Voom.res, function(x) names(x[1:200])),
                 dd=lapply(DD.res, function(x) names(x[1:200])),
                 mast=lapply(MAST.res, function(x) names(x[1:200])),
                 exc_cepo=features_celltype[[1]],
                 exc_voom=features_celltype[[2]],
                 exc_dd=features_celltype[[3]],
                 exc_mast=features_celltype[[4]])

all.list.collapse <- lapply(all.list, function(x) {
  unique(unlist(x))
})


counter <- 0
lapply(all.list.collapse[5:8], function(x) {
  counter <<- counter + 1
  set_n <- c("cepo", "voom", "dd", "mast")[[counter]]
  pdf(file=paste0("../../../Figures/trajectory_analysis/trajectory_dx_upset_exluded_", set_n,".pdf"), width=6, height=6)
  print(UpSetR::upset(UpSetR::fromList(list(Cepo=all.list.collapse[[1]],
                                            Voom=all.list.collapse[[2]],
                                            DD=all.list.collapse[[3]],
                                            MAST=all.list.collapse[[4]],
                                            excluded=x
  )), nsets=5))
  dev.off()
})

```

# Generate trajectories from refined gene sets
```{r}
list_full_model = tibble(
  features = features_final,
  features_name = names(features),
  dyno_data = purrr::map(
    .x = features, 
    .f = ~ make_dyno(sce[.x,], prop = 1L)),
  method = "mst",
  model = purrr::map2(
    .x = dyno_data, 
    .y = method,
    .f = ~ infer_trajectory(
      dataset = .x, 
      method = ti_mst(dimred = "mds")) %>% 
      add_dimred(dyndimred::dimred_mds, 
                 expression_source = .x$expression)),
  celltype_plot = purrr::map2(
    .x = model,
    .y = dyno_data,
    .f = ~ plot_dimred(trajectory = .x, 
                       expression_source = .y$expression, 
                       grouping = .y$cell_info$group))
)

list_full_model

save(list_full_model, file="../../../Data/trajecotories_HSC_traj_dx_excl_results.RData")
```

# Visualize trajectories
```{r fig.height=5, fig.width=20}
patchwork::wrap_plots(list_full_model$celltype_plot, 
                      ncol = 4, nrow=1) %>% ggsave(filename="../../../Figures/trajectory_analysis/trajectory_dx_trajectories_exclusion.pdf", width=10, height=3)
```

# Session info
```{r}
sessioninfo::session_info()
```
