---
title: "Trajectory analysis of HSPC differentiation using refined gene sets"
author: "Kevin Wang and Hani Kim"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This script runs multiple trajectory analysis on the HSPC differentiation dataset using [dynverse](https://dynverse.org/) framework. After features have been identified from running our benchmarking differential analysis algorithms.

# Loading 

## Packages
```{r, message = FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(SingleCellExperiment)
  library(scater)
  library(hdf5r)
  library(dyno)
  library(tidyverse)
  library(dyntoy)
  library(dyneval)
  library(patchwork)
  library(furrr)
})

#dynwrap::test_docker_installation(detailed = TRUE)
plan(multisession, workers = 10)
```

## Load HSPC single cell data and Dx results
```{r}
load("../../../Data/Dx_benchmark_result_HSPC.RData", verbose = T)
load("../../../Data/trajectory_sce_HSC.RData", verbose=T)
```

## Source functions
```{r}
source("../../../R/utils.R")
```

# Feature set
```{r}
list_dx_results = list(
    Cepo = DS.res, 
    Voom = Voom.res, 
    DD = DD.res,
    MAST = MAST.res)

extract_top_genes = function(dx_result, n){
    unique(unlist(lapply(dx_result, function(x) names(x)[1:n])))
}
```

## Define root cell
```{r}
set.seed(2020)
root <- sample(colnames(sce[sce$group %in% "Stem"]), 1)
```

# Single test case using `infer_trajectories

```{r, eval = FALSE}
dyno_tmp = make_dyno(sce[head(features1, 100),], prop = 1L)
dyno_data = add_prior_information(dataset = dyno_tmp, start_id = root, start_n = 1, end_n = 2)

model = infer_trajectory(dataset = dyno_data, method = ti_mst(dimred = "mds")) %>%
    add_dimred(dyndimred::dimred_mds, expression_source = dyno_data$expression)

model_withoutmds = infer_trajectory(dataset = dyno_data, method = ti_mst()) %>%
    add_dimred(dyndimred::dimred_mds, expression_source = dyno_data$expression)

plot_dimred(model, dimred = "mds", grouping = dyno_data$cell_info$group)
plot_dimred(model_withoutmds, dimred = "mds", grouping = dyno_data$cell_info$group)

dimred <- dyndimred::dimred_umap(dyno_data$expression)
plot_dimred(model, dimred = dimred, grouping = dyno_data$cell_info$group)
plot_dimred(model_withoutmds, dimred = dimred, grouping = dyno_data$cell_info$group)
```

# Setting up grid of parameters before computations

```{r}
param_grid = tidyr::expand_grid(
    list_dx_results,
    n_features = seq(40,200,20)) %>% 
    dplyr::mutate(
        features_name = paste0(names(list_dx_results), "__",
                               n_features),
        features = purrr::map2(.x = list_dx_results,
                              .y = n_features,
                              .f = ~ extract_top_genes(dx_result = .x, n = .y)),
        dyno_data = purrr::map(
            .x = features,
            .f = ~ make_dyno(sce[.x,], prop = 1L) %>%
                add_prior_information(dataset = ., start_id = root, start_n = 1, end_n = 2))
    )

param_grid
```

# Trajectory methods 
```{r}
method_list = list(
    mst = ti_mst(dimred = "mds"), 
    slingshot = ti_slingshot(),
    projected_monocle = ti_projected_monocle(),
    monocle_ddrtree = ti_monocle_ddrtree(),
    celltree_maptpx = ti_celltree_maptpx(),
    slice = ti_slice(),
    celltree_vem = ti_celltree_vem(),
    monocle_ica = ti_monocle_ica()
)
```

# Compute model and consolidate into a grid
```{r}
list_models = purrr::map(
    .x = param_grid$dyno_data,
    .f = ~ infer_trajectories(dataset = .x, method = method_list, map_fun = furrr::future_map))

result_grid = param_grid %>% 
    dplyr::mutate(list_models) %>% 
    unnest(list_models)

result_grid
```

# Save result 
```{r}
save(result_grid, file = "../../../Data/trajecotories_HSC_traj_dx_results_all_methods.RData")
```

# Session info
```{r}
sessioninfo::session_info()
```
