---
title: "Rare cell type analysis on Cellbench data"
author: "Hani Kim"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This script runs the benchmarking differential analysis algorithms on Cellbench data (3 cell types) where the cell types are iteratively subsampled to emulate rare cell types. A total of 5% of the cells from a given cell types are subsampled. The subsampling is performed 50 times. 

# Loading 

## Packages
```{r, message = FALSE}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(MAST)
  library(limma)
  library(edgeR)
})
```

## Load 3-class cellbench data
```{r}
load("../../../Data/cellbench_sce_3class.RData", verbose=T)
rm(sce_CELseq2_log)
rm(sce_Dropseq_log)
```

## Source functions
```{r}
source("../../../R/Cepo.R") 
source("../../../R/doDE.R")
source("../../../R/doDD.R")
source("../../../R/utils.R")
source("../../../R/doSample.R")
```

# Extract cell type labels
```{r}
##### generate cell type labels
cty <- factor(sce_10x_log$celltype)
```

# Run Dx on subsampled data
```{r}
system.time(results_DS <- doSample(sce_10x_log, method="Cepo", n=50, p=0.05))
names(results_DS) <- levels(cty)

system.time(results_DD <- doSample(sce_10x_log, method="DD", n=50, p=0.05))
names(results_DD) <- levels(cty)

system.time(results_MAST <- doSample(sce_10x_log, method="MAST", n=50, p=0.05))
names(results_MAST) <- levels(cty)

system.time(results_Voom <- doSample(sce_10x_log, method="Voom", n=50, p=0.05))
names(results_Voom) <- levels(cty)

save(results_DS, results_DD, results_MAST, results_Voom, file="../../../Data/cellbench_rare_celltype_five_percent.RData")
```

# Session Info
```{r}
sessioninfo::session_info()
```


