---
title: "Cellbench data processing"
author: "Hani Kim"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
Cellbench data were downloaded from [sc_mixcology](https://github.com/LuyiTian/sc_mixology). This script performs filtering of doublets using the doublet labels from the authors and cleans up the data for use in our manuscript.   

# Loading 

## Packages
```{r, message = FALSE}
library(biomaRt)
```

# Process 3-class cellbench data
```{r}
load("../../../Data/sincell_with_class.RData", verbose=T)
```

Remove doublets and extract cell type labels
```{r}
sce_10x_log <- sce_sc_10x_qc[, sce_sc_10x_qc$demuxlet_cls == "SNG"] 
sce_10x_log$celltype <- sce_10x_log$cell_line_demuxlet 

sce_CELseq2_log <- sce_sc_CELseq2_qc[, sce_sc_CELseq2_qc$demuxlet_cls == "SNG"] 
sce_CELseq2_log$celltype <- sce_CELseq2_log$cell_line_demuxlet 

sce_Dropseq_log <- sce_sc_Dropseq_qc[, sce_sc_Dropseq_qc$demuxlet_cls == "SNG"]
sce_Dropseq_log$celltype <- sce_Dropseq_log$cell_line_demuxlet 
```

Get overlapping genes between datasets, convert rownames and remove any duplicate/missing rows
```{r}
idx <- Reduce(intersect, list(rownames(sce_10x_log),
                              rownames(sce_CELseq2_log),
                              rownames(sce_Dropseq_log)))

ensembl <- useMart("ensembl", dataset="hsapiens_gene_ensembl")
human_gene_ids <- idx
genes <- getBM(attributes=c('ensembl_gene_id',
                            'external_gene_name'),
               values = human_gene_ids,
               mart = ensembl)
geneID <- genes$external_gene_name
names(geneID) <- genes$ensembl_gene_id

quickProcess <- function(mat) {
    mat <- mat[idx,]
    rownames(mat) <- geneID[rownames(mat)]
    mat <- mat[!is.na(rownames(mat)),]
    mat <- mat[!duplicated(rownames(mat)),]
    return(mat)
}

sce_10x_log <- quickProcess(sce_10x_log)
sce_CELseq2_log <- quickProcess(sce_CELseq2_log)
sce_Dropseq_log <- quickProcess(sce_Dropseq_log)
```

Save data
```{r}
save(sce_10x_log, sce_CELseq2_log, sce_Dropseq_log, file="../../../Data/cellbench_sce_3class.RData")
```

# Process 5-class cellbench data
```{r}
load("../../../Data/sincell_with_class_5cl.RData", verbose=T)
```

Repeat processing as above
```{r}
sce_10x_5cl_log <- sce_sc_10x_5cl_qc[, sce_sc_10x_5cl_qc$demuxlet_cls == "SNG"]
sce_10x_5cl_log$celltype <- sce_10x_5cl_log$cell_line_demuxlet

quickProcess <- function(mat) {
    mat <- mat[!is.na(rownames(mat)),]
    mat <- mat[!duplicated(rownames(mat)),]
    return(mat)
}

sce_10x_5cl_log <- quickProcess(sce_10x_5cl_log)

save(sce_10x_5cl_log, file="../../../Data/cellbench_sce_5class.RData")
```

# Session Info
```{r}
sessioninfo::session_info()
```